Comparing summed probability distributions of shoreline and radiocarbon dates from the Mesolithic Skagerrak coast of Norway

### 2023-06-30 analysis/paper/paper.Rmd
Added additional dates from Steinar and made map of 14C-dated sites

# Map prepared in 01_shoreline_spd.R 
knitr::include_graphics(here("analysis/figures/map1.png"))
```

@@

```{r sumshore, echo = FALSE, fig.cap = "A) The distribution of the shoreline dated sites (n = 921) by elevation. The binwidth is 4 m. B) The mean of \\textit{terminus post quem} dates for the sites, plotted with a binsize of 200 years, reflecting the common approach to summing shoreline dates. C) The result of instead summing the probability of the shoreline dates achieved with the method employed here, which accounts for the likely distance between sites and shoreline. The binwidth is 5 years, following the resolution used with the dating procedure.",  out.width = "500px", fig.align="center", regfloat = TRUE}

@@

```{r spdmc, echo = FALSE, fig.cap = "Monte Carlo simulation for shoreline and radiocarbon dates. A) Summed probability of shoreline dated sites (n = 921) compared to an exponential null model. B) SSPD compared to a logistic null model. C) SSPD compared to a uniform null model. D) Summed probability of calibrated radiocarbon dates (n = 310, bins = 134) compared to an exponential null model. E) RSPD compared to a logistic null model. F) RSPD compared to a uniform null model.", out.width = "500px", fig.align="center", regfloat = TRUE}
knitr::include_graphics(here("analysis/figures/mc.png"))
```
For the RSPD, the logistic model achieved a p-value of, and could therefore not be rejected. Based on the available data, this therefore represents the best explanation of the data. For the SSPD all of the standard models could be rejected.

@@

While it appears that the SSPD and RSPD are differentially impacted by population numbers, and, we suggest, the SSPD is more heavily influenced by land-use and mobility patterns, this suggestion and the nature of the discrepancy between the proxies clearly requires further investigation. Some lines of inquiry could be to also draw on lithic technology and tool-kit diversity as a function of population size and density [e.g. @collard2013a; @riede2014; @solheim2020], and quantitative measures for mobility patterns that can potentially be inferred from the composition of lithic assemblages [e.g. @roalkvam2022]. Furthermore, environmental conditions, constraints and developments also represent a range of factors that can have implications for these dimensions, and will consequently be of future importance to unpack their relationship [e.g. @ordonez2022; @jorgensen2020].

>>>>

# Map prepared in 01_shoreline_spd.R 
knitr::include_graphics(here("analysis/figures/map2.png"))
```

@@

```{r sumshore, echo = FALSE, fig.cap = "A) The distribution of the shoreline dated sites (n = 921) by altitude above present sea-level. The binwidth is 4 m. B) The mean of \\textit{terminus post quem} dates for the sites, plotted with a binsize of 200 years, reflecting the common approach to summing shoreline dates. C) The result of instead summing the probability of the shoreline dates achieved with the method employed here, which accounts for the likely distance between sites and shoreline. The binwidth is 5 years, following the resolution used with the dating procedure.",  out.width = "500px", fig.align="center", regfloat = TRUE}

@@

```{r spdmc, echo = FALSE, fig.cap = "Monte Carlo simulation for shoreline and radiocarbon dates. A) Summed probability of shoreline dated sites (n = 921) compared to an exponential null model. B) SSPD compared to a logistic null model. C) SSPD compared to a uniform null model. D) Summed probability of calibrated radiocarbon dates (n = 310, bins = 134) compared to an exponential null model. E) RSPD compared to a logistic null model. F) RSPD compared to a uniform null model.", out.width = "500px", fig.align="center", regfloat = TRUE}
knitr::include_graphics(here("analysis/figures/mc.png"))
```
For the RSPD, the logistic model achieved a p-value of 0.108, and could therefore not be rejected. Based on the available data, this therefore represents the best explanation of the data. For the SSPD all of the standard models could be rejected.

@@

While it appears that the SSPD and RSPD are differentially impacted by population numbers, and, we suggest, the SSPD is more heavily influenced by land-use and mobility patterns, this suggestion and the nature of the discrepancy between the proxies clearly requires further investigation. Some lines of inquiry could be to also draw on lithic technology and tool-kit diversity as a function of population size and density [e.g. @collard2013a; @riede2014; @solheim2020], and mobility patterns that can potentially be inferred from the composition of lithic assemblages [e.g. @barton2014; @clark2017; @preston2018; @roalkvam2022]. Furthermore, environmental conditions, constraints and developments also represent a range of factors that can have implications for these dimensions, and will consequently be of future importance to unpack their relationship [e.g. @jorgensen2020; @ordonez2022].
